---
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, results='hide',message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(weathermetrics)
library(ThermIndex)

dat<- read_csv("/work/jessebell/puvvula/h_def_2021nov.csv")

dat$date <- as.Date(dat$date, format = "%m/%d/%Y")
dat$month<- as.factor(month(dat$date))
dat$year<- as.factor(format(dat$date, '%Y'))
dat$dow <- wday(as.Date(dat$date, format = "%m/%d/%Y"))
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dat$wDay <- factor((weekdays(dat$date) %in% weekdays1), 
                       levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

fct<- c(2, 13:68)
dat[,fct]<-lapply(dat[,fct],factor)

dat$tmax_cel<- as.numeric(tempftoc(dat$tmax)) #convert F to C
dat$HI_17<- as.factor(ifelse(dat$tmax_cel >= 35, 1,0))


#filter by region

coas<- dat %>% subset(Region=="Coastal")
pied<- dat %>% subset(Region=="Piedmont")

#additional definitions for Coastal
coas$shw_17 <- as.factor(ifelse(coas$tmax_cel >= 35, 1,0))
coas$shw_26 <- as.factor(ifelse(coas$app_temp >= 95.85, 1,0)) #85th pct
coas$shw_27 <- as.factor(ifelse(coas$app_temp >= 97.16, 1,0)) #90th pct
coas$shw_28 <- as.factor(ifelse(coas$app_temp >= 98.99, 1,0)) #95th pct


#additional definitions for Piedmont
quantile(pied$app_temp, probs = c(0.95, 0.90, 0.85))

pied$shw_17 <- as.factor(ifelse(pied$tmax_cel >= 35, 1,0))
pied$shw_26 <- as.factor(ifelse(pied$app_temp >= 95.47, 1,0)) #85th pct
pied$shw_27 <- as.factor(ifelse(pied$app_temp >= 96.66, 1,0)) #90th pct
pied$shw_28 <- as.factor(ifelse(pied$app_temp >= 98.52, 1,0)) #95th pct

coas$pop<-as.numeric(2741101)
pied$pop<- as.numeric(5571983)
#Coastal: 2741101
#Piedmont: 5571983

```
```


#model comparision - Coastal region

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Coastal
library(MASS)

c1<- glm.nb(imp_cnt~as.factor(shw_15)+wDay+month+year+NWS_HW+offset(log(pop)),data=coas)
c2<- glm.nb(imp_cnt~as.factor(shw_08)+wDay+month+year+NWS_HW+offset(log(pop)),data=coas)
c3<- glm.nb(imp_cnt~as.factor(shw_07)+wDay+month+year+NWS_HW+offset(log(pop)),data=coas)
c4<- glm.nb(imp_cnt~as.factor(shw_16)+wDay+month+year+NWS_HW+offset(log(pop)),data=coas)
c5<- glm.nb(imp_cnt~as.factor(mhw_07)+wDay+month+year+NWS_HW+offset(log(pop)),data=coas)

anova(c1,c2,c3,c4,c5, test="Chisq")

```












#model comparision - Piedmont region

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Piedmont

p1<- glm.nb(imp_cnt~as.factor(shw_08)+wDay+month+year+NWS_HW+offset(log(pop)),data=pied)
p2<- glm.nb(imp_cnt~as.factor(shw_07)+wDay+month+year+NWS_HW+offset(log(pop)),data=pied)
p3<- glm.nb(imp_cnt~as.factor(shw_15)+wDay+month+year+NWS_HW+offset(log(pop)),data=pied)
p4<- glm.nb(imp_cnt~as.factor(mhw_07)+wDay+month+year+NWS_HW+offset(log(pop)),data=pied)
p5<- glm.nb(imp_cnt~as.factor(mhw_28)+wDay+month+year+NWS_HW+offset(log(pop)),data=pied)

anova(p1,p2,p3,p4,p5, test="Chisq")

```

